---
title: "Lobster Report for Biol 806"
author: 'Shea, Muhammad, and Adriel'
format: html
---

## Data Source

All data used in this report is sourced from Reed and Miller (2022). Full citation: 

Santa Barbara Coastal LTER, D. Reed, and R. Miller. 2022. SBC LTER: Reef: Abundance, size and fishing effort for California Spiny Lobster (Panulirus interruptus), ongoing since 2012 ver 8. Environmental Data Initiative. https://doi.org/10.6073/pasta/25aa371650a671bafad64dd25a39ee18 (Accessed 2026-04-23).

Data consists of abundance, size, and commercial fishing impact on California spiny lobsters. This information is spread across two separate csv files.

## Analysis and Visualizations by Shea (repo owner)

```{r}
#| output: false
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(here)

lobster_abundance <- read_csv(here::here("data/Lobster_Abundance_All_Years_20220829.csv")) %>% 
    mutate(SIZE_MM = na_if(SIZE_MM, -99999))

lobsters_summarize <- lobster_abundance %>% 
  group_by(SITE, YEAR) %>% 
  summarize(COUNT = n())
```

```{r}
ggplot(data = lobsters_summarize, aes(x = YEAR, y = COUNT)) +
  geom_point(aes(color = SITE)) +
  geom_line(aes(color = SITE)) +
  labs(y = 'Lobster count',
       x = 'Year',
       title = 'Lobster Abundance over time') +
  theme_classic()
```

## Analysis and Visualizations by Muhammad (collaborator)

```{r}
#| output: false
lobster_traps <- read_csv(here::here(
  "data/Lobster_Trap_Counts_All_Years_20210519.csv"
)) %>% 
  mutate(TRAPS = na_if(TRAPS, -99999))

lobsters_traps_summarize <- lobster_traps %>% 
  group_by(SITE, YEAR) %>% 
  summarize(TOTAL_TRAPS = sum(TRAPS, na.rm = TRUE), .groups = "drop")
```

```{r}
ggplot(
  data = lobsters_traps_summarize,
  aes(x = YEAR, y = TOTAL_TRAPS, color = SITE)
) +
  geom_point(size = 3, alpha = 0.8) +
  geom_line(linewidth = 1) +
  labs(
    title = "Total Lobster Traps by Site Over Time",
    x = "Year",
    y = "Total Number of Traps",
    color = "Site"
  ) +
  theme_light() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "right"
  )
```

## Discussion

It looks like Lobster abundance might have peaked between 2017 and 2020, driven mainly at the AQUE site. Some sites are less impacted by traps than others. There may be a trend toward decreasing trap numbers over time at some sites (AQUE, MOHK, and ABUR), however the dearth of traps in the latest sample is likely due to a lack of data.
