---
title: "Lobster Report for Biol 806"
author: 'Shea, Muhammad, and Adriel'
format: html
---

## Data Source

All data used in this report is sourced from Reed and Miller (2022). Full citation: 

Santa Barbara Coastal LTER, D. Reed, and R. Miller. 2022. SBC LTER: Reef: Abundance, size and fishing effort for California Spiny Lobster (Panulirus interruptus), ongoing since 2012 ver 8. Environmental Data Initiative. https://doi.org/10.6073/pasta/25aa371650a671bafad64dd25a39ee18 (Accessed 2026-04-23).

Data consists of abundance, size, and commercial fishing impact on California spiny lobsters. This information is spread across two separate csv files.

## Analysis and Visualizations by Shea (repo owner)

```{r}
#| output: false
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(here)

lobster_abundance <- read_csv(here::here("data/Lobster_Abundance_All_Years_20220829.csv")) %>% 
    mutate(SIZE_MM = na_if(SIZE_MM, -99999))

lobsters_summarize <- lobster_abundance %>% 
  group_by(SITE, YEAR) %>% 
  summarize(COUNT = n())
```

```{r}
ggplot(data = lobsters_summarize, aes(x = YEAR, y = COUNT)) +
  geom_point(aes(color = SITE)) +
  geom_line(aes(color = SITE)) +
  labs(y = 'Lobster count',
       x = 'Year') +
  theme_classic()
```

## Discussion

It looks like Lobster catches peaked between 2017 and 2020, driven mainly at the AQUE site.
